Haematology II

### 2018-05-04 haem2.rmd
sc

@@

>>>>

Mech of fever: exogenous pyrogens => IL-1 and TNF => increased cyclooxygenase activity => $\Uparrow$ PG including PGE~2~ => reset temperature set point in hypothalamus

Fatigue in children + *infn*

-  Acute: *URTI*, sinus *infn*, hepatitis, chickenpox, measles, meningitis
-  More prolonged: *mononucleosis*, CMV, OSA
-  Chronic: tb, diabetes, *IBD*, CKD, thyroid disease, heart failure, poor nutrition, *cancer*, abscesses

Anaemia needs to be severe to cause persistent fatigue in children

"Growing pains": shouldn't affect the chest (sternum) but can affect children as young as 5

If generalised bone pain => really should consider leukaemia

Pale during fever because of vasodilation and less blood going to skin

Ix:

-  Good idea to do uric acid, LDH, Ca, Phos as a baseline in case tumour lysis syndrome

@@

Amoxiciliin + clavulanic acid: common for strep

-  EBV
-  HIV
-  CMV
-  Acute hepatitis: abdo symptoms
-  Lymphoma: drenching night sweats
-  Withdrawal symptoms from alc/drugs can cause night sweats
-  Infective endocarditis

Histoplasmosis can come from bat bites

Pharyngitis is mostly viral

First line for tonsilitis is penicillin (amoxiclav)

Ix:

-  FBC
-  ELFT + creatinine = liver + kidney
-  CXR: tb in lungs
-  USS: spleen size, but it won't be diagnostic of prognostic so a bit useless..

Bloods:

-  Dehydration: osmolality high and urea high
-  ALT and AST: ALT = hepatocellular, AST = alcohol (sometimes)
-  Liver impairment
